---
title: "First_Knitr"
author: "EHarris"
date: "5/25/2017"
output: html_document
---

## Set Global Option
```{r setoptions, echo = FALSE}
opts_chunk$set(echo = FALSE, results = “as”, include = TRUE)
```

## Load anticipated packages to be used throughout Course 5 Project 1
```{r setup, include=TRUE}
ipak <- function(pkg){
      new.pkg <- pkg[!(pkg %in% installed.packages()[, "Package"])]
      if (length(new.pkg)) 
            install.packages(new.pkg, dependencies = TRUE)
      sapply(pkg, require, character.only = TRUE)
}

# usage
packages <- c("plyr", "dplyr", "data.table", "dtplyr", "lubridate", "ggplot2", "knitr")
ipak(packages)
```

## 1. Code for reading in the dataset and/or processing the data
### Step 1: Read csv file into R
```{r fileload, include = TRUE}
activity <- read.csv("activity.csv", header = TRUE)
```

### Step 2a: Process data to address other parts of assignment
```{r iniital_processing, include = TRUE}
activity$date1 <- as.Date(activity$date, format = "%Y-%m-%d")
activity$day <- weekdays(activity$date1)
activity$date1 <- format(activity$date1, format = "%Y%m%d")
activity$date2 <- format(activity$date1, format = "%Y%b%d")
activity <- select(activity, date, date1, date2, day, interval, steps)
weekday <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")
weekend <- c("Saturday", "Sunday")
activity.wday <- subset(activity, day %in% weekday)
activity.wend <- subset(activity, day %in% weekend)
```

### Step 2b: Review content / layout of dataset
```{r fileview, include = TRUE}
dim(activity)

summary(activity)

table(activity$date, useNA = "always")
```

dates <- data.frame(date=unique(activity$date1))
daterange <- c(dates)
min.date <- dates$date[1]
max.date <- dates[length(dates),]


### Total Number of steps take each day
```{r total_steps_per_day, include = TRUE}
steps.a.day <- aggregate(steps ~ date1, data = activity, FUN = function(x) sum=sum(x))
```

### Mean & Median steps per day [Days with no steps are excluded]
```{r mean_median_steps_per_day, include = TRUE}
meansteps <- mean(steps.a.day$steps, na.rm = TRUE)
mediansteps <- median(steps.a.day$steps, na.rm = TRUE)
steps.a.day1 <- data.frame("Mean Steps" = meansteps, "Median Steps" = mediansteps)
steps.a.day1
```

### Average number of steps taken per 5 minute interval
```
This can be viewed two ways:  
1. Mean number of steps in a 5-minute interval
2. Mean number of steps for each 5-minute interval(same interval of each day)  
```

#### Mean number of steps in a 5-minute interval
```{r mean_steps_per5, include = TRUE}
steps.per.5min <- aggregate(steps ~ date1 + interval, data = activity,
                                    FUN = function(x) meansteps=mean(x))
steps.per.5min <- steps.per.5min[order(steps.per.5min[,1], steps.per.5min[,2]),]
mean.5min <- mean(steps.per.5min$steps)
mean.5min
```

### Mean number of steps each 5-minute interval
```{r mean_steps_each5, include = TRUE}
steps.each.5min <- aggregate(steps ~ interval, data = activity,
                                    FUN = function(x) meansteps=mean(x))
steps.each.5min
```

### 2. Histogram of the total number of steps taken each day.  Two steps:  
###    a) Summarize data to support requested histogram information  
###    b) Plot data to histogram
```{r totalsteps}
stepaday <- aggregate(steps ~ date, data = activity, FUN = function(x) sum=sum(x))
mrgtotal <- merge(dates[,1], stepaday[,1:2], by = 1, all = TRUE)
```

### Plot histogram reflecting total number of steps by day
```{r steps_per_dzy_hist, fig.keep = "all"", fig.show = "asis" }
with(steps.a.day, plot(date1, steps, main = "Total Steps Each Day", type = "n"))
qplot(date, steps, data = mrgtotal, xlim = c(mrgtotal[1]), geom = "histogram",
            main = "Total Steps Each Day", xlab = "Dates", ylab = "# Steps")
```

### Panel plot of average number of steps per 5-minute interval  
#### Comparing weekdays and weekend
```{r panel_plot, include = TRUE}
par(mfrow = c(1, 2), mar = c(4, 4, 2, 1), oma = c(0, 0, 2, 0))
with(airqualit, {
      plot(Wind, Ozone, main = “Ozone and Wind”)
	plot(Solar.R, Ozone, main = “Ozone and Solar Radiation”)

